---
title: "Fit the biofilm colonisation area with a stationary solution from the model"
author: "Andres Diaz,Tatyana Pichugina, Paul Rainey"
date: "6/13/2019"
output:
  github_document
  #pdf_document:
  #      toc: true
  #      toc_depth: 2

#pdf_document
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggridges)
library(gridExtra)
library(stringr)
library(modelr)
library(broom)
library(ggpubr)
source("../../src/LinearRegresionFit.R")
source("../../src/LinearRegresionFit_perLayer.R")



```


## Integral area
Integral area represent sum of the colonized pixel for each frame. Each frame represent one hour of experiment. Here we are interested in 15 first hours.


```{r read and filter data, echo=FALSE, message=FALSE, warning=FALSE}
# read file 
AreaY<-read_delim(file='../../data/ResultTable_SliceY.txt',col_names = TRUE,delim=',')

# extract frame and strain type from the name 
AreaY<-AreaY %>% 
  # new colums with strain
  separate(Experiment,c('Strain'),sep='_',remove=FALSE,extra="drop") %>%
  mutate(Strain=as.factor(Strain) )%>%
  # extract data frame
  mutate(frame=as.numeric(str_split(datafile,'_',simplify = TRUE)[,5]),data_stamp=as.numeric(str_split(datafile,'_',simplify = TRUE)[,3]))

# take 15 hours of experiment
AreaY<-AreaY %>% filter(frame %in% seq(from=1,to=15,by=1)) # frame 1-15

```


```{r introduce Integral area, echo=FALSE, message=FALSE, warning=FALSE}

# Integral area
IntegralArea<-AreaY %>%
  group_by(Strain,Experiment,data_stamp,frame)  %>%
  summarise(IntAreaTotal=sum(SumSliceY),IntAreaGain=sum(SumSliceYgain),
            IntAreaLost=sum(SumSliceYlost),IntAreaSame=sum(SumSliceYsame)) %>%
  mutate(GainTotal=IntAreaGain/IntAreaTotal,LostTotal=IntAreaLost/IntAreaTotal,
         SameTotal=IntAreaSame/IntAreaTotal) %>%
  filter(-IntAreaGain,-IntAreaLost,-IntAreaSame) %>%
  gather(key=Area_type,value='Area',IntAreaTotal,GainTotal,LostTotal,SameTotal,factor_key = TRUE)


IntegralArea %>% filter(Area_type=='IntAreaTotal') %>% 
  ggplot(aes(x=frame,y=Area,color=Strain))+
  geom_line(aes(group=data_stamp),size=2)+
  ggtitle('Integral area') + 
  xlab("time frame in hours") + ylab("Area")+
  theme_bw(base_size = 12)
```


##  Integlal area ALI and GALI

We cutted each image to the two parts: ALI part includes 200px (32mkm) layer counted from the ALI, and  GLI part includes rest part of the image. 
The integral area per Strain expreriment is shown below.


```{r echo=FALSE, message=FALSE, warning=FALSE}
IntegralArea_ALI<-AreaY %>%
  filter(Layer<400 & frame>5) %>%
  group_by(Strain,Experiment,data_stamp,frame) %>%
  summarise(IntAreaTotal=sum(SumSliceY)) 

PlotALI<-IntegralArea_ALI %>% ggplot(aes(x=frame,y=IntAreaTotal,color=Strain))+
    geom_line(aes(group=data_stamp),size=2)+
    ggtitle('Integral area ALI') + 
    xlab("time frame in hours") + ylab("Area")+
    theme_bw(base_size = 12)

IntegralArea_GLI<-AreaY %>%
  filter(Layer>200 & frame>10) %>%
  group_by(Strain,Experiment,data_stamp,frame) %>%
  summarise(IntAreaTotal=sum(SumSliceY)) 

PlotGLI<-IntegralArea_GLI %>% ggplot(aes(x=frame,y=IntAreaTotal,color=Strain))+
    geom_line(aes(group=data_stamp),size=2)+
    ggtitle('Integral area GLI') + 
    xlab("time frame in minutes") + ylab("Area")+
    theme_bw(base_size = 12)

ggarrange(PlotALI,PlotGLI,ncol=2, nrow =1,common.legend = TRUE, legend="bottom")

```

We fitted ALI and GALI separately by Log(Area)~A+B*frame. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

fit_lm<-function(df) lm(log(IntAreaTotal)~frame,data=df)

FitCoeff_GLI<-LinearRegressionFit(IntegralArea_GLI,"Integral Area GLI")
FitCoeff_GLI<-FitCoeff_GLI %>% mutate(type="GLI")
FitCoeff_ALI<-LinearRegressionFit(IntegralArea_ALI,"Integral Area ALI")
FitCoeff_ALI<-FitCoeff_ALI %>% mutate(type="ALI")

FitCoeff<-bind_rows(FitCoeff_GLI,FitCoeff_ALI) %>% mutate(Itype=as.factor(type))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
FitCoeff %>% filter(term=="frame") %>%
    ggplot(aes(x=Strain,y=estimate, fill=type))+
    geom_boxplot()+
    scale_fill_manual(values=c("#b2df8a", "lightblue"))+
    geom_hline(yintercept=0, linetype="dashed", color = "black", size=1)+
    ylab("colonisation rate 1/min")+
    xlab("Strain")+
    labs(title="Colonization rate for ALI and GLI")+
    theme_grey(base_size = 12)

```
## Colonisation area per Layer 

LayerSize=32 mkm (200px)

```{r fig.width=6, fig.height=8,echo=FALSE,message=FALSE, warning=FALSE,}
ggplot(AreaY,aes(x=frame,y=SumSliceY,color=Strain))+
  geom_line(aes(group=Experiment),size=1)+facet_grid(Layer~.)+
  theme_bw(base_size = 10)+
  labs(title="Area per layer from ALI")+
  ylab("Area per Layer")
```
### Occupation proportion
```{r echo=FALSE, message=FALSE, warning=FALSE,}
AreaY %>% group_by(Experiment,frame) %>% mutate(SumYTotal=sum(SumSliceY)) %>%
  mutate(SumSliceYProp=ifelse(SumYTotal>0,SumSliceY/SumYTotal,0)) %>%
  ggplot(aes(x=frame,y=SumSliceYProp,color=as.factor(Layer)))+
  geom_point(aes(group=data_stamp),size=2)+facet_wrap(~Experiment)+
  theme_bw()
```

### Colonisation area per Layer 
```{r fig.width=6, fig.height=8,echo=FALSE,message=FALSE, warning=FALSE,}
nest_by_layer<-AreaY %>% filter(frame>5) %>% group_by(Layer) %>%
          nest()

#first data frame
i=1
CurrentLayer<-nest_by_layer$Layer[i]
FitCoeffLayer<-LinearRegresionFit_perLayer(nest_by_layer$data[[i]],as.character(CurrentLayer))
FitCoeffLayerTotal<-FitCoeffLayer %>% mutate(Layer=CurrentLayer)

for(i in 2:nrow(nest_by_layer)){
  CurrentLayer<-nest_by_layer$Layer[i]
  FitCoeffLayer<-LinearRegresionFit_perLayer(nest_by_layer$data[[i]],as.character(CurrentLayer))
  FitCoeffLayer <-FitCoeffLayer %>% mutate(Layer=CurrentLayer)
  FitCoeffLayerTotal<-bind_rows(FitCoeffLayerTotal,FitCoeffLayer) 
}
```

```{r fig.width=8, fig.height=6,echo=FALSE,message=FALSE, warning=FALSE,}
A<-FitCoeffLayerTotal %>% filter(term=="frame") %>% 
  ggplot(aes(x=Strain.x,y=estimate, fill=Strain.x))+
  geom_boxplot()+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.1)+
  geom_jitter(shape=16, position=position_jitter(0.2))+
  geom_hline(yintercept=0, linetype="dashed", color = "darkgrey", size=1)+
  ylab("colonisation rate 1/min")+
  xlab("Layer")+
  labs(title="Colonization rate per Layer")+
  theme_classic(base_size = 10)+
  theme(axis.text.x = element_blank(),axis.ticks=element_blank())+
  facet_grid(cols =vars(Layer), scales = "fixed",switch="x")
print(A)
```

### Quality of the fit: R^2 per experiment
```{r fig.width=10, fig.height=6, echo=FALSE,message=FALSE, warning=FALSE,}
A<-FitCoeffLayerTotal %>% filter(term=="frame") %>% 
  ggplot(aes(x=Experiment,y=r.squared,fill=Strain.x,label = Experiment))+
  geom_bar(stat="identity", position=position_dodge())+
  geom_errorbar(aes(ymin=r.squared-std.error, ymax=r.squared+std.error), width=.2,
                  position=position_dodge(.9))+
  ylab("r.squared")+
  xlab("Layer")+
  labs(title="R^2")+
  coord_flip()+
  theme_bw(base_size = 8)+
  facet_grid(cols =vars(Layer), scales = "fixed")
print(A)
```  
```{r fig.width=10, fig.height=6, echo=FALSE,message=FALSE, warning=FALSE,}
A<-FitCoeffLayerTotal %>% filter(term=="(Intercept)") %>% 
  ggplot(aes(x=Experiment,y=estimate,fill=Strain.x,label = Experiment))+
  geom_bar(stat="identity", position=position_dodge())+
  geom_errorbar(aes(ymin=r.squared-std.error, ymax=r.squared+std.error), width=.2,
                  position=position_dodge(.9))+
  ylab("intersept")+
  xlab("Layer")+
  labs(title="Intersept")+
  coord_flip()+
  theme_bw(base_size = 8)+
  facet_grid(cols =vars(Layer), scales = "fixed")
print(A)
```  


<!-- #### Plot shows ratio  between Area per Layer to total colonization area for different time frames -->
<!-- ```{r,fig.height=14, fig.width=12,echo=FALSE,message=FALSE, warning=FALSE} -->
<!-- AreaTotal<-AreaY %>% group_by(data_stamp,frame,Strain) %>% -->
<!--     summarise(AreaTotal=sum(SumSliceY)) -->

<!-- AreaYSumSliceY<-AreaY %>% select(data_stamp,Strain,Layer,frame,SumSliceY) %>% -->
<!--   inner_join(AreaTotal,by=c("data_stamp","frame","Strain")) %>%  -->
<!--   filter(AreaTotal!=0) %>%  -->
<!--   mutate(Ratio=SumSliceY/AreaTotal)  -->


<!-- AreaYSumSliceY %>% filter(Strain=='dwss') %>% -->
<!--   ggplot(aes(x=Layer,y=Ratio),color=as.numeric(Layer))+geom_bar(size=1,stat="identity",fill="darkblue")+ -->
<!--   scale_x_reverse()+ -->
<!--   coord_flip()+ -->
<!--   facet_grid(frame~data_stamp)+ -->
<!--   ggtitle('dwss ratio')+ -->
<!--   theme_gray(base_size = 10) -->

<!-- AreaYSumSliceY %>% filter(Strain=='SM') %>% -->
<!--   ggplot(aes(x=Layer,y=Ratio),color=as.numeric(Layer))+geom_bar(size=1,stat="identity",fill="darkblue")+ -->
<!--   scale_x_reverse()+ -->
<!--   coord_flip()+ -->
<!--   facet_grid(frame~data_stamp)+ -->
<!--   ggtitle('SM ratio')+ -->
<!--   theme_gray(base_size = 10) -->
<!-- ``` -->



