---
title: "Fit the biofilm colonisation area with a stationary solution from the model"
author: "Andres Diaz,Tatyana Pichugina, Paul Rainey"
date: "7/02/2019"
output:
  github_document
  #pdf_document:
  #      toc: true
  #      toc_depth: 2

#pdf_document
 
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggridges)
library(gridExtra)
library(stringr)
library(modelr)
library(broom)
library(ggpubr)

source("../../src/LinearRegresionFit.R")
source("../../src/LinearRegresionFit_perLayer.R")

```

```{r Load Data, echo=FALSE, message=FALSE, warning=FALSE}
# read file from mpistaff drive
Data<-read_delim(file='/Volumes/mpistaff/Diaz_Pichugina_Pseudomona/Data/1-TIMELAPSES_2019_1-1__Results/AllData06_20_2019_461-TIMELAPSES_2019_1-1__Results/SM_dwss_all.csv',col_names = TRUE,delim=',')

# one pixel per mkm
# add column with data stamp
# extract frame and strain type from the name
Data<-Data %>% 
  # new colums with strain
  mutate(strain=as.factor(strain) )%>%
  mutate(data_stamp=ifelse(strain=="SM",substr(FILE,16,28),substr(FILE,16,30))) %>%
  # filter probe
  filter((X!=0 & Y!=9.5)) 

# rename column with complex name
colnames(Data)[str_detect(colnames(Data),"Surface")]<-"SurfaceMkm"
colnames(Data)[colnames(Data)=="Nb of obj. voxels"]<-"SurfacePx"
OnePixelinMkm<-Data$SurfaceMkm[1]/Data$SurfacePx[1]*10

#Bounding box: smallest box encompassing the object: if this box is ticked,
#the coordinates of the upper left corner of the box, its width, height and depth
#are retrieved and logged
colnames(Data)[colnames(Data)=="B-width"]<-"BoxWidth"
colnames(Data)[colnames(Data)=="B-height"]<-"BoxHight"

# reduce number of colums
DataF<-Data %>% select(data_stamp,strain,interval,SurfacePx,X,Y,BX,BY,BoxWidth,BoxHight)

# take 15 hours of experiment, filter probe
DataF<-DataF %>% filter(interval %in% seq(from=3,to=15,by=1))
```



## Total integral area
Integral area represent sum of the colonized pixel for each frame. Each frame represent one hour of experiment. Here we are interested in 15 first hours.


```{r Total Integral Area, echo=FALSE, message=FALSE, warning=FALSE,fig.cap = "Total_Integral_Area"}
DataSumInt<-DataF %>% group_by(strain,data_stamp,interval) %>%
  summarise(IntAreaTotal=sum(SurfacePx)) %>%
  # transfer to the mkm^2
  mutate(IntAreaTotal=IntAreaTotal*OnePixelinMkm*OnePixelinMkm)

ggplot(DataSumInt,aes(x=interval,y=IntAreaTotal,group=data_stamp,color=strain)) +
  geom_line(size=1)+
  scale_color_manual(values = c("dwss" = "#e41a1c", "SM" = "#377eb8"))+
  ggtitle('Total integral area')+
  xlab("time in hour") + ylab("Colonized area mkm^2")+
  theme_bw(base_size = 12)

```
## Total integral area was fitted by Log(Area)~A+B*time

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.cap = "Integral_Area_Fit_results"}
DataSumInt <- DataSumInt %>% rename(frame=interval,Experiment=data_stamp,Strain=strain) 

fit_lm<-function(df) lm(log(IntAreaTotal)~frame,data=df)
FitCoeff<-LinearRegressionFit(DataSumInt,"Integral Colonisation Area")

```

```{r Ali position, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# binning data on x axis (8 layers), choose point with min Bounding box BY
Data_Binned<-DataF %>% mutate(BX_binned=cut(BX, 8,left=FALSE,right=TRUE)) %>%
            group_by(data_stamp,interval,BX_binned) %>%
            summarise(ALI=min(BY)) 

# binning data 
Data_Binned %>% ggplot(aes(x=BX_binned,y=ALI,color=interval))+
  geom_point()+facet_wrap(~data_stamp)+scale_y_reverse()+theme(axis.text.x = element_text(angle = 90, hjust = 1))

# estimate position as 0.01 quantile
EstimateALI<-Data_Binned %>% group_by(data_stamp) %>% summarise(QALI=quantile(ALI,0.01))

# compare with min edge of bounding box histogramme
Data_Binned<-left_join(Data_Binned,EstimateALI)
Data_Binned %>% filter(ALI<500) %>% ggplot(aes(x=ALI))+
  geom_histogram()+geom_vline(aes(xintercept = QALI),color="red")+
facet_wrap(~data_stamp)
            
# add to the filtered data
DataF<-left_join(DataF,EstimateALI)
DataF <- DataF %>% mutate(Y=Y-QALI)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
## Check how good the ALI position was estimated
#library(magick)
#images_folder<-'/Volumes/mpistaff/Diaz_Pichugina_Pseudomona/Data/1-TIMELAPSES_2019_1-1__Results/Test_ALI_position/'
#list_experiments<-unique(DataF$data_stamp)
#list_experiments_file<-map(list_experiments,paste,"_FR_PJ.tif",sep="")
#list_experiments_output<-map(list_experiments,paste,"_FR_PJ_ALI.avi",sep="")

#for(i in seq(1:length(list_experiments))){
#    print(i)
#    file_name<-paste(images_folder,list_experiments_file[i],sep="")
#    ALI=round(filter(DataF,data_stamp==list_experiments[i])$QALI[1])

#    # Produce image  with ALI-line 
#    Image<-image_read(file_name) %>%
#          image_annotate("experiment",gravity='northwest',color = "white") %>%
#          image_draw()
#          abline(h=ALI,col="red",lwd=5)
#          dev.off()
#    # Write to the file
#    file_name_output<-paste(images_folder,list_experiments_output[i])  
 #   image_write(Image,path=file_name_output)  
#}
```
## Colonisation area per Layer 
LayerSize=64 mkm (400px)

```{r echo=FALSE,message=FALSE, warning=FALSE,fig.cap = "Area_per_Layer"}
DataLayer<-DataF %>% mutate(Y=Y*OnePixelinMkm) %>% mutate(Layer=cut(Y, seq(0,2048*OnePixelinMkm,400*OnePixelinMkm),right = TRUE,left=TRUE)) %>% 
                    filter(Layer!="NA") %>% group_by(strain,data_stamp,interval,Layer) %>%
                    summarise(SumSliceY=sum(SurfacePx))

ggplot(DataLayer,aes(x=interval,y=SumSliceY*OnePixelinMkm*OnePixelinMkm,group=data_stamp,color=strain))+facet_grid(Layer~.)+
  scale_color_manual(values = c("dwss" = "#e41a1c", "SM" = "#377eb8"))+
  geom_line(size=1)+
  ggtitle('Colonized area per Layer')+
  xlab("time in hour") + ylab("Colonized area in mkm^2")+
  theme_bw(base_size = 12)+
  theme_bw()

```
## Linear regression fit separetely for each layer
```{r fig.width=6, fig.height=8,echo=FALSE,message=FALSE, warning=FALSE,fig.cap = "Area_per_Layer_Fit_Results"}
DataLayer<-DataLayer %>% rename(frame=interval,Experiment=data_stamp,Strain=strain) 
nest_by_layer<-DataLayer %>% group_by(Layer) %>%
          nest()

#first data frame
i=1
CurrentLayer<-nest_by_layer$Layer[i]
FitCoeffLayer<-LinearRegresionFit_perLayer(nest_by_layer$data[[i]],as.character(CurrentLayer))
FitCoeffLayerTotal<-FitCoeffLayer %>% mutate(Layer=CurrentLayer)

for(i in 2:nrow(nest_by_layer)){
  CurrentLayer<-nest_by_layer$Layer[i]
  FitCoeffLayer<-LinearRegresionFit_perLayer(nest_by_layer$data[[i]],as.character(CurrentLayer))
  FitCoeffLayer <-FitCoeffLayer %>% mutate(Layer=CurrentLayer)
  FitCoeffLayerTotal<-bind_rows(FitCoeffLayerTotal,FitCoeffLayer) 
}
```

```{r fig.width=8, fig.height=6,echo=FALSE,message=FALSE, warning=FALSE,fig.cap = "Area_per_Layer_Fit_Results"}
FitCoeffLayerTotal %>% filter(term=="frame") %>% 
  ggplot(aes(x=Strain.x,y=estimate, fill=Strain.x))+
  geom_boxplot()+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.1)+
  geom_jitter(shape=16, position=position_jitter(0.2))+
  geom_hline(yintercept=0, linetype="dashed", color = "darkgrey", size=1)+
  ylab("colonisation rate 1/min")+
  xlab("Layer in mkm")+
  labs(title="Colonization rate per Layer")+
  theme_classic(base_size = 10)+
  theme(axis.text.x = element_blank(),axis.ticks=element_blank())+
  scale_fill_manual(values = c("dwss" = "#e41a1c", "SM" = "#377eb8"))+
  facet_grid(cols =vars(Layer), scales = "fixed",switch="x")

```
## Quality of the fit per Layer: R^2 per experiment
```{r fig.width=10, fig.height=6, echo=FALSE,message=FALSE, warning=FALSE,fig.cap = "Area_per_Layer_Fit_Results_R2"}
A<-FitCoeffLayerTotal %>% filter(term=="frame") %>% 
  ggplot(aes(x=Experiment,y=r.squared,fill=Strain.x,label = Experiment))+
  geom_bar(stat="identity", position=position_dodge())+
  geom_errorbar(aes(ymin=r.squared-std.error, ymax=r.squared+std.error), width=.2,
                  position=position_dodge(.9))+
  scale_fill_manual(values = c("dwss" = "#e41a1c", "SM" = "#377eb8"))+
  ylab("r.squared")+
  xlab("Layer")+
  labs(title="R^2")+
  coord_flip()+
  theme_bw(base_size = 8)+
  facet_grid(cols =vars(Layer), scales = "fixed")
print(A)
```
## Quality of the fit per Layer: Intersept per experiment
```{r fig.width=10, fig.height=6, echo=FALSE,message=FALSE, warning=FALSE,fig_cap="Area_per_Layer_Fit_Results_Intersept"}
A<-FitCoeffLayerTotal %>% filter(term=="(Intercept)") %>% 
  ggplot(aes(x=Experiment,y=estimate,fill=Strain.x,label = Experiment))+
  geom_bar(stat="identity", position=position_dodge())+
  geom_errorbar(aes(ymin=r.squared-std.error, ymax=r.squared+std.error), width=.2,
                  position=position_dodge(.9))+
  scale_fill_manual(values = c("dwss" = "#e41a1c", "SM" = "#377eb8"))+
  ylab("intersept")+
  xlab("Layer")+
  labs(title="Intersept")+
  coord_flip()+
  theme_bw(base_size = 8)+
  facet_grid(cols =vars(Layer), scales = "fixed")
print(A)
```  
## Occupation proportion per experiment
```{r echo=FALSE, message=FALSE, warning=FALSE,fig.cap="Occupancy_percent_Layer"}
# create color palette
library(RColorBrewer)
myColors <-c("#b2182b","#d6604d","#92c5de","#4393c3","#2166ac")
names(myColors) <- levels(DataLayer$Layer)


DataLayer %>% filter(Strain=="dwss") %>% group_by(Experiment,frame) %>% mutate(SumYTotal=sum(SumSliceY)) %>%
  mutate(SumSliceYProp=ifelse(SumYTotal>0,SumSliceY/SumYTotal,0)) %>%
  ggplot(aes(x=frame,y=SumSliceYProp,color=as.factor(Layer)))+
  geom_line(aes(group=Layer),size=1)+facet_wrap(~Experiment)+
  scale_colour_manual(name = "Layer",values = myColors)+
  labs(title="Occupancy percent per Layer")+
  ylab("Occupancy percent")+
  xlab("time in hour")+
  theme_bw()


DataLayer %>% filter(Strain=="SM") %>% group_by(Experiment,frame) %>% mutate(SumYTotal=sum(SumSliceY)) %>%
  mutate(SumSliceYProp=ifelse(SumYTotal>0,SumSliceY/SumYTotal,0)) %>%
  ggplot(aes(x=frame,y=SumSliceYProp,color=as.factor(Layer)))+
  geom_line(aes(group=Layer),size=1)+facet_wrap(~Experiment)+
  labs(title="Occupancy percent per Layer")+
  scale_colour_manual(name = "Layer",values = myColors)+
  ylab("Occupancy percent")+
  xlab("time in hour")+
  theme_bw()



```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE,echo=FALSE}
#DataF %>% filter(Y>1000 & Y<2000) %>% ggplot(aes(x=interval,y=SurfacePx,group=data_stamp))+facet_wrap(~data_stamp,scales="free")+
#  geom_line(size=1)

#for (i in unique(DataF$data_stamp)) {
#  PlotHist<-DataLayer %>% filter(data_stamp== i) %>% ggplot(aes(x=interval,y=SurfacePx, group=interval, #fill=strain))+
#  geom_violin()+facet_wrap(Ybinned~data_stamp,scales="free")
#  print(PlotHist)
}
```







